---
title: 'DATA 607 - Assignment 7'
author: "Mia Chen"
date: "March 17, 2019"
output: html_document
---

#Working with HTML, XML and JSON in R

* Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.

* Take the information that you've selected about these three books, and separately create three files which
store the book's information in HTML (using an html table), XML, and JSON formats (e.g. "books.html",
"books.xml", and "books.json"). To help you better understand the different file structures, I'd prefer that you
create each of these files "by hand" unless you're already very comfortable with the file formats.

* Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?

* Your deliverable is the three source files and the R code. If you can, package your assignment solution up into
an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files
accessible from the web].

```{r message=FALSE}
#load packages
library(XML)
library(RJSONIO)
library(dplyr)
library(DT)
```

***

#HTML 


    <!DOCTYPE html>
    <html>
    	<title>Favortie Books in Art History</title>
    	<body>
    		<table>
    			<tr>
    				<th>Title</th>
    				<th>Author</th>
    				<th>Edition</th>
    				<th>Year</th>
    				<th>ISBN</th>
    			</tr>
    			<tr>
    				<td>Art History Volume 1</td>
    				<td>
    					<div>Marilyn Stokstad</div>
    					<div>Michael W. Cothren</div>
    				</td>
    				<td>Fifth</td>
    				<td>2013</td>
    				<td>978-0205873487</td>
    			</tr>
    			<tr>
    				<td>Ancient Greece: From Prehistoric to Hellenistic Times</td>
    				<td>Thomas R. Martin</td>
    				<td>Second</td>
    				<td>2013</td>
    				<td>978-0300160055</td>
    			</tr>
    			<tr>
    				<td>Ancient Rome: From Romulus to Justinian</td>
    				<td>Thomas R. Martin</td>
    				<td>First</td>
    				<td>2013</td>
    				<td>978-0300198317</td>
    			</tr>
    		</table>
    
    	</body>
    </html>


To make the text file accessible from the web, I uploaded the html file to github
```{r message=FALSE}
#read url from github
html_url <- "https://raw.githubusercontent.com/miachen410/DATA607/master/books.html"

#download as local file
download.file(html_url, destfile = "~/books.html")

#read from local file using htmlParse()
books_html <- file.path("books.html") %>% htmlParse()

#read html table
html_table <- readHTMLTable(books_html)

#convert to a data frame with proper header
html_df <- html_table[[1]] %>% as.data.frame()
```

```{r}
#view as table
datatable(html_df)
```

***

#XML


    <?xml version="1.0" encoding="UTF-8"?>
    <books>
    	<book>
    		<Title>Art History Volume 1</Title>
    		<Author>Marilyn Stokstad, Michael W. Cothren</Author>
    		<Edition>Fifth</Edition>
    		<Year>2013</Year>
    		<ISBN>978-0205873487</ISBN>
    	</book>
    	<book>
    		<Title>Ancient Greece: From Prehistoric to Hellenistic Times</Title>
    		<Author>Thomas R. Martin</Author>
    		<Edition>Second</Edition>
    		<Year>2013</Year>
    		<ISBN>978-0300160055</ISBN>
    	</book>
    	<book>
    		<Title>Ancient Rome: From Romulus to Justinian</Title>		
    		<Author>Thomas R. Martin</Author>		
    		<Edition>First</Edition>
    		<Year>2013</Year>
    		<ISBN>978-0300198317</ISBN>
    	</book>
    </books>


To make the text file accessible from the web, I uploaded the xml file to github
```{r message=FALSE}
#read url from github
xml_url <- "https://raw.githubusercontent.com/miachen410/DATA607/master/books.xml"

#download as local file
download.file(xml_url, destfile = "~/books.xml")

#read XML file with xmlParse() to obtain the XML tree with R
books_xml <- file.path("books.xml") %>% xmlParse(file = "books.xml")

#convert to a data frame
xml_df <- xmlToDataFrame(books_xml)
```

```{r}
#view as a table
datatable(xml_df)
```

***

#JSON

    {"books" :[
  	  {
    		"Title" : "Art History Volume 1",
    		"Author" : ["Marilyn Stokstad", Michael W. Cothren"],
    		"Edition" : "Fifth",
    		"Year" : 2013,
    		"ISBN" : "978-0205873487"
    	},
    	{
    		"Title" : "Ancient Greece: From Prehistoric to Hellenistic Times",
    		"Author" : "Thomas R. Martin",
    		"Edition" : "Second",
    		"Year" : 2013,
    		"ISBN" : "978-0300160055"
    	},
    	{
    		"Title" : "Ancient Rome: From Romulus to Justinian",
    		"Author" : "Thomas R. Martin",
    		"Edition" : "First",
    		"Year" : 2013,
    		"ISBN" : "978-0300198317"
    	}
    ]}

    
To make the text file accessible from the web, I uploaded the json file to github
```{r message=FALSE}
#read url from github
json_url <- "https://raw.githubusercontent.com/miachen410/DATA607/master/books.json"

#download as local file
download.file(json_url, destfile = "~/books.json")

#read from local file
json_file <- file.path("books.json")

#check if JSON file is valid
isValidJSON(json_file)
```

```{r}
#read content in JSON format and converts it to R objects
books_json <- fromJSON(content = json_file)

#check class to see if it's a list object
class(books_json)
```

```{r}
#use lapply() function to turn the lists into data frames
#join the data frames with a do.call() on rbind()
json_df <- do.call("rbind", lapply(books_json[[1]], data.frame))

#display the data frame in a data table
datatable(json_df)
```

***

> ##Observation:
> HTML and XML data frames are idential. However, the JSON data frame has an extra row - a separated entry for each of the authors.
