---
title: "DATA607 - Project 2"
author: "Mia Chen"
date: "March 10, 2019"
output: html_document
---

*** 

###**Goal of this assignment**: 

####- Practice in preparing different datasets for downstream analysis work.

Your task is to choose any three of the "wide" datasets identified in the Week 6 Discussion items.For each of the three chosen datasets:

* Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a "wide" structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.

* Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

* Perform the analysis requested in the discussion item.

* The code is in an R Markdown file, posted to rpubs.com, which includes narrative descriptions of data cleanup work, analysis, and conclusions.

***

Retail Volume & Avocado Price for 2018
 
The table below represents weekly 2018 retail scan data for National retail volume (units) and price. Retail scan data comes directly from retailers' cash registers based on actual retail sales of Hass avocados. Starting in 2013, the table below reflects an expanded, multi-outlet retail data set. Multi-outlet reporting includes an aggregation of the following channels: grocery, mass, club, drug, dollar and military. The Average Price (of avocados) in the table reflects a per unit (per avocado) cost, even when multiple units (avocados) are sold in bags. The Product Lookup codes (PLU's) in the table are only for Hass avocados. Other varieties of avocados (e.g. greenskins) are not included in this table.

source: http://www.hassavocadoboard.com/retail/volume-and-price-data

***

Load packages
```{r message=FALSE}
library(tidyr)
library(dplyr)
library(lubridate)
```

Read csv file into R
```{r}
avocado <- read.csv("https://raw.githubusercontent.com/miachen410/DATA607/master/avocado.csv")
```

Take a look at the first 6 rows of the dataset
```{r}
head(avocado)
```

Take a look at the first 6 rows of the dataset
```{r}
tail(avocado)
```

Let's look at the structure of the dataset
```{r}
str(avocado)
```

Let's look at the summary of the dataset
```{r}
summary(avocado)
```

###Data tidying and transformation

#### Gather columns(Part 1)
- Columns X4046, X4225 X4770 should be gathered under the variable "ProductCode"
- We will use the gather function in tidyr to turn the wide table into long table: Product Code and Volume show as key-value pairs
```{r}
avocado2 <- avocado %>% gather(ProductCode, Volume, -Date:-Total.Volume, -Total.Bags:-XLarge.Bags)
# New column ProductCode, with values gather from X4046, X4225 X4770, excluding all other columns
head(avocado2)
```

#### Gather columns(Part 2)
- Columns Small.Bags, Large.Bags, XLarge.Bages should be gathered under the variable "BagSize"
- Again, we will use the gather function in tidyr
```{r}
avocado3 <- avocado2 %>% gather(BagSize, Bags, -Date:-Total.Bags, -ProductCode, -Volume)
head(avocado3)
```

Now we look at the structure again
```{r}
str(avocado3)
```

Change the class of Date from "factor" to "Date" using lubridate package
```{r}
avocado3$Date <- ymd(avocado3$Date)
class(avocado3$Date)
```

Let's look at the distribution of avocado average price:
- The average price is centered around 1.0 to 1.1
```{r}
hist(avocado3$AveragePrice)
```

Relationship between total volume and average price showing on a scatterplot:
- There is a decreasing linear relationship: the more the cheaper
```{r}
plot(avocado3$Total.Volume, avocado3$AveragePrice)
```

Average Price changes throughout the year:
- Higher price in January, September and December
```{r}
avocado3 %>% select(Date, AveragePrice) %>% plot()
```

Let's find out the price differences among the products:
```{r}
avocado3 %>% group_by(Month = month(Date)) %>% summarize(TotalVolume = sum(Volume), AvgPrice = mean(AveragePrice)) %>% arrange(desc(AvgPrice))
# For each month, summarize the sum of volume and mean of average price, arrange from high price to low price
```

```{r}
avocado3 %>% group_by(Month = month(Date)) %>% summarize(TotalVolume = sum(Volume), AvgPrice = mean(AveragePrice)) %>% arrange(desc(TotalVolume))
# For each month, summarize the sum of volume and mean of average price, arrange from high quantity to low quantity
```

####Conclusion: December 2018 is the month which avocado had the highest price and lowest volume.

