---
title: "DATA607 Assignment 2 - R and SQL"
author: "MiaC"
date: "February 7, 2019"
output: html_document
---

For this assignment, I choose six recent popular movies and asked at least five people that I know to rate each of these movie that they have seen on a scale of 1 to 5. I took the results and stored them in a MySQL database connected to AWS. Here, I'm going to load the information into an R dataframe.

```{r message = FALSE}
library(RMySQL)
library(DBI)
library(DT)
library(getPass)
```

Connect to movie database in MySQL through AWS
```{r}
con <- dbConnect(RMySQL::MySQL(), 
  dbname = "movie",
  host = "mydb.ccjuueb9wecv.us-east-2.rds.amazonaws.com",
  port = 3306,
  user = "Admin", 
  password = getPass())
```

Retrieve table names in the database
```{r}
dbListTables(con)
```

Retrieve data from the movies table
```{r}
movies <- dbReadTable(con, "movies")
datatable(movies)
```

Retrieve data from the reviews table
```{r}
reviews <- dbReadTable(con, "reviews")
datatable(reviews)
```

Query and join MySQL tables
```{r}
movie_ratings_query <- dbGetQuery(con, "SELECT m.movie_name, r.rating 
           FROM movies AS m
           LEFT JOIN reviews AS r
           ON m.movie_id = r.movie_id")
```

Create data frame from the joined query
```{r}
movie_ratings_df <- as.data.frame(movie_ratings_query)
datatable(movie_ratings_df)
```

